Wait-free Queues with Polygarithmic Step Complexity

### 2022-07-27 article/psuedocode.tex
review dsearch

\begin{lemma}[DSearch correctness] \label{dsearch}
Assume \nf{root.blocks[end].num\sub{enq}}$\geq$\nf{e} and $E_{root, e}$ is the response to some \nf{Dequeue()} in \nf{root.blocks[end]}.  \nf{DSearch(e, end)} returns \nf{<b, i>} such $E_{root,b,i}= E_{root,e}$ in $\Theta \big(\log($\nf{root.blocks[b].\size + root.blocks[end].\size} \hspace{-0.8em}$)\big)$ steps.
\end{lemma}
\begin{proof}
First we show \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. From line \ref{addOP}, we know that size of the every block in the tree is greater than 0. So each block in \nf{root.blocks[b..end]} contains at least one \nf{Enqueue()} or at least one \nf{Dequeue()}. There cannot be more than \nf{root.blocks[b].\size} \nf{Dequeue()}s in \nf{root.blocks[b+1..end-1]}, since the  elements of the queue after \nf{blocks[b]} would become dequeued at some point, after \nf{root.blocks[b]} operations finish end before \nf{root.blocks[end]} operations start. And the response to to some \nf{dequeue()} in \nf{root.blocks[end]} could not be in $E(n,b)$. Furthermore since the size of the queue would become \nf{root.blocks[end].\size} in the end and there cannot be more than \nf{root.blocks[b].\size} \nf{Dequeue()}s so there cannot be more than \nf{root.blocks[b].\size + root.blocks[end].\size} \nf{enqueue()} operations in \nf{root.blocks[b..end]}. Cause if there were more, then the \nf{root.blocks[end]}'s size would become more than \nf{root.blocks[end].\size}. Overall there can be at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} operations in \nf{root.blocks[b+1..end-1]} and since each block size is $\geq 1$ thus there are at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]}. So \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. See Figure \ref{end-b}.

Now that we know there are at most \nf{root.blocks[b].\size+root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]} then with doubling search in $\Theta \big(\log($\nf{root.blocks[b].\size+root.blocks[end].\size}$)\big)$ steps we reach \nf{start=c} that the \nf{root.blocks[c].sum\sub{enq}} is less than \nf{e} and \nf{end-c} is not more than $2\times$\nf{root.blocks[b].\size+root.blocks[end].\size}. Cause if otherwise, then \nf{(end-c)/2} satisfied the \nf{root.blocks[(end-c)/2].sum\sub{enq}<e}. Note that in the line \ref{doubling} the distance between \nf{end} and \nf{stat} is doubled. See figure \ref{fig::doubling}.

After computing \nf{b}, the value \nf{i} is computed via the definition of \nf{sum\sub{enq}} in constant time (Line \ref{DSearchComputei}). So the routine non constant part is the binary search which takes $\Theta(\log$\nf{root.blocks[b].\size+root.blocks[end].\size)}$)$ steps from the first paragraph.
\end{proof}
\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/doubling.png}

>>>>

\begin{lemma}[DSearch correctness] \label{dsearch}
Assume \nf{root.blocks[end].sum\sub{enq}}$\geq$\nf{e} and $E_{root, e}$'s \nf{element} is the response to some \nf{Dequeue()} operation in \nf{root.blocks[end]}.  \nf{DSearch(e, end)} returns \nf{<b, i>} such that $E_{root,b,i}= E_{root,e}$.
\end{lemma}
\begin{proof}
It is trivial to see that the doubling search from \nf{root.blocks[end]} to \nf{root.blocks[0]} will find $E_{root,e}$ eventually. Because \nf{root.blocks[].sum\sub{enq}} is an increasing value from 0 to some value greater than \nf{e}. So there is a \nf{b} that \nf{root.blocks[b].sum\sub{enq}}$>e$ but \nf{root.blocks[b-1].sum\sub{enq}}$<e$.

First we show \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. From line \ref{addOP}, we know that size of the every block in the tree is greater than 0. So each block in \nf{root.blocks[b..end]} contains at least one \nf{Enqueue} or at least one \nf{Dequeue}. Suppose there were more than \nf{root.blocks[b].\size}\nf{Dequeue}s in \nf{root.blocks[b+1..end-1]}. Then the  queue would become empty at some point after \nf{blocks[b]}'s last operations and before \nf{root.blocks[end]}'s first operation. Which means the response to to a \nf{Dequeue} in \nf{root.blocks[end]} could not be in $E(n,b)$. Furthermore since the size of the queue would become \nf{root.blocks[end].\size} after the \nf{root.blocks[end]}, there cannot be more than \nf{root.blocks[b].\size + root.blocks[end].\size} \nf{Enqueue}s. Because there can be at most \nf{root.blocks[b].\size}\nf{Dequeue}s and the final size is \nf{root.blocks[end].\size}. Overall there can be at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} operations in \nf{root.blocks[b+1..end-1]} and since each block size is $\geq 1$ thus there are at most $2 \times$\nf{root.blocks[b].\size+ root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]}. So \nf{end-b}$\leq 2 \times$\nf{root.blocks[b].\size+root.blocks[end].\size+1}. See Figure \ref{end-b}.

Now that we know there are at most \nf{root.blocks[b].\size+root.blocks[end].\size} blocks in between \nf{root.blocks[b]} and \nf{root.blocks[end]} then with doubling search in $\Theta \big(\log($\nf{root.blocks[b].\size+root.blocks[end].\size}$)\big)$ steps we reach \nf{start=c} that the \nf{root.blocks[c].sum\sub{enq}} is less than \nf{e} and \nf{end-c} is not more than $2\times$\nf{root.blocks[b].\size+root.blocks[end].\size}. Beause otherwise, then \nf{(end-c)/2} satisfied the \nf{root.blocks[(end-c)/2].sum\sub{enq}<e}. In line \ref{doubling} the differnece between \nf{end} and \nf{start} is doubled. See Figure \ref{fig::doubling}.

 After computing \nf{b}, the value \nf{i} is computed via the definition of \nf{sum\sub{enq}} in constant time (Line \ref{DSearchComputei}). So the routine non constant part is the binary search which takes $\Theta(\log$\nf{root.blocks[b].\size+root.blocks[end].\size)}$)$ steps from the first paragraph.

\end{proof}
\begin{figure}[hbt]  
  \center\includegraphics[width=6in]{pics/doubling.png}
